---
title: "Mixed Models"
output: html_document
---

Analysis of genetic (co)variation within and among populations

# Setup and Import Data

```{r}
library(tidyverse)
library(lme4)
```


```{r}
InDat<-read.csv("./data/FloralDisplayInd.csv")
```


# Check data

```{r}
hist(InDat$Start)
hist(InDat$Duration)
hist(InDat$Peak)
hist(InDat$Var)
hist(InDat$Skew)
hist(InDat$Kurt)

InDat[InDat$Var > 5000 & !is.na(InDat$Kurt),c("Num","Fam","Pop","Var")]# Outlier?
InDat[InDat$Skew <= -4 & !is.na(InDat$Kurt),c("Num","Fam","Pop","Skew")]# Outlier?
InDat[InDat$Kurt >= 10 & !is.na(InDat$Kurt),c("Num","Fam","Pop","Kurt")]# Outlier?

```

> Note to Dale: Check the data for this individual


# Scale Data

## Key parameters

1. Days to first flower
2. Duration of flowering 
2. Flowering schedule variance 
3. Days to peak flowering


Scale traits to mean of zero and unit variance

```{r}
PDat<-InDat %>%
  select(Num:Lat,Start,Peak,Duration,Var,Skew,Kurt) %>%
  mutate(S_Start = (Start-mean(Start,na.rm=T))/sd(Start,na.rm=T)) %>%
  mutate(S_Peak = (Peak-mean(Peak,na.rm=T))/sd(Peak,na.rm=T)) %>%  
  mutate(S_Duration = (Duration-mean(Duration,na.rm=T))/sd(Duration,na.rm=T)) %>%  
  mutate(S_Var = (Var-mean(Var,na.rm=T))/sd(Var,na.rm=T)) %>%
  mutate(S_Skew = (Skew-mean(Skew,na.rm=T))/sd(Skew,na.rm=T)) %>%
  mutate(S_Kurt = (Kurt-mean(Kurt,na.rm=T))/sd(Kurt,na.rm=T))  
```

# Reorganize

Move from 'wide' to 'long' format for mixed models analysis

```{r}
MDat<-PDat %>%
  select(Num:Lat, S_Start, S_Peak, S_Duration, S_Var, S_Skew, S_Kurt) %>%
  gather(key = "Fparm", value = "Value",
       S_Start, S_Peak, S_Duration, S_Var, S_Skew, S_Kurt)
MDat$PseudoGroup<-as.factor("Group")
```

# Statistics

Is there detectable genetic variation for these parameters?

**PROBLEM**: Not enough within-family replication
(Most families have only 1 or 2 individuals)

See README.md for ideas

## Which phenology traits differ among populations ?

```{r}
  Mod0 <- lmer(Peak ~ 1 + (1|Fam), data = InDat)
  Mod1 <- lmer(Peak ~ 1 + Pop + (1|Fam), data = InDat)


for (trait in unique(MDat$Fparm)){
  Mod0 <- lmer(Value ~ 1 + (1|Fam), data = MDat[ MDat$Fparm == trait , ])
  Mod1 <- lmer(Value ~ 1 + Pop + (1|Fam), data = MDat[ MDat$Fparm == trait , ])
  print("--------------")
  print(trait)
  print(anova(Mod0,Mod1))
  print("--------------")
  
  Mod0<-Mod1<-NA # Clean-up
}
```


## Which traits form clines?

```{r}
for (trait in unique(MDat$Fparm)){
  Mod0 <- lmer(Value ~ 1 + (1|Pop), data = MDat[ MDat$Fparm == trait , ])
  Mod1 <- lmer(Value ~ 1 + Lat + (1|Pop), data = MDat[ MDat$Fparm == trait , ])
  print("--------------")
  print(trait)
  print(anova(Mod0,Mod1))
  print("--------------")
  
  Mod0<-Mod1<-NA # Clean-up
}
```

## Do populations differ after accounting for clines?

```{r}
for (trait in unique(MDat$Fparm)){
  Mod0 <- lmer(Value ~ 1 + Lat + (1|Pop), data = MDat[ MDat$Fparm == trait , ])
  Mod1 <- lmer(Value ~ 1 + Lat + Pop + (1|Pop/Fam), data = MDat[ MDat$Fparm == trait , ])
  print("--------------")
  print(trait)
  print(anova(Mod0,Mod1))
  print("--------------")
  
    Mod0<-Mod1<-NA # Clean-up

}
```

> NOTE: only compare for significant clines




