---
title: "Floral Display Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyr)
library(viridis)
library(lme4)
```

# Data import & setup

```{r}
Disp<-read.csv("LythrumDisplay2008.csv")
Disp$Block<-as.factor(gsub("[a-f]","",Disp$Row))
Disp$RelFit<-Disp$HinfW/mean(Disp$HinfW,na.rm=T)
head(Disp)
```

## Data description

  * **Num** -- unique number from 1 to `r max(Disp$Num)`
  * **Row** -- combination of block (1-4) and row within block (a-f)
  * **Pop** -- Population code (see [Montague et al. 2008](https://onlinelibrary.wiley.com/doi/full/10.1111/j.1420-9101.2007.01456.x))
  * **Lat** -- Latitude of seed family origin
  * **Start** -- Julian Day of first flower (standardized to 0 for first observed flower)
  * **End** -- Julian Day of final flower
  * **Veg** -- Vegetative height at first flower
  * **PC1** -- PC1 of Vegetative height and date of first flower
  * **HinfW** -- Inflorescence Biomass at harvest
  * **TotFlwr** -- Total # of flowers produced
  * **d0 to d60** -- Number of viable flowers on for each 5-day period, beginning with first day of flower observed
  * **Block** -- Experimental block (1-4)
  * **RelFit** -- Relative fitness (calculated from inflorescence biomass)

## Rearrage and organize

```{r}
Days<-names(Disp)[grep("d[0-9]",names(Disp))] # Vector of names
Flwr<-Disp %>% gather(Days,key="RelDay",value="OpFlwr") # Rearrange
Flwr$RelDay<-as.numeric(gsub("d","",Flwr$RelDay)) # Remove 'd' from day
```

## Add floral display metrics

```{r}
Flwr$AbsDay<-Flwr$RelDay+Flwr$Start # Add absolute day
Flwr$CumFlwr<-ave(Flwr$OpFlwr,Flwr$Num,FUN=cumsum) # Calculate cumulative N flowers
```

### Standardize metrics

Standardize floral display as a proportion of total flowers to control for variation in plant size (we already know there are latitudinal clines in size).

```{r}
FlwrStd<-function(x){x/max(x,na.rm=T)}
FlwrStd2<-function(x){x/sum(x,na.rm=T)}
Flwr$StdCumFlwr<-ave(Flwr$CumFlwr,Flwr$Num,FUN=FlwrStd)
Flwr$StdOpFlwr<-ave(Flwr$OpFlwr,Flwr$Num,FUN=FlwrStd2)
```

# Inspect flowering profiles

## Open flowers from each population

```{r}
ggplot(aes(x=RelDay,y=StdOpFlwr,group=Pop,colour=Lat),data=Flwr[Flwr$StdOpFlwr>0 & Flwr$StdCumFlwr<1,])+
  geom_point(alpha=I(0.3))+stat_smooth(method="lm",se=F,alpha=I(0.3),formula=y~x+I(x^2))+
  theme_bw()+
  scale_color_gradientn(colours=c("#FE5F55","#9D8DF1","#71A9F7","#BEE9E8"))+
  labs(x="Day (relative)",y="Relative # flowers")
```

## Cumulative flowering for each plant

```{r}
ggplot(aes(x=RelDay,y=StdCumFlwr,group=Num,colour=Lat),data=Flwr)+
  geom_point(alpha=I(0.3))+geom_line(alpha=I(0.3))+
  theme_bw()+
  scale_color_gradientn(colours=c("#FE5F55","#9D8DF1","#71A9F7","#BEE9E8"))
```

# Is log-log relationship approximately linear?
```{r}
ggplot(aes(x=log(RelDay),y=log(StdCumFlwr),group=Num,colour=Lat),data=Flwr[Flwr$StdOpFlwr>0 & Flwr$StdOpFlwr<1,])+
  geom_point(alpha=I(0.3))+stat_smooth(geom="line",method="lm",se=F,alpha=0.3)+
  theme_bw()+
  scale_color_gradientn(colours=c("#FE5F55","#9D8DF1","#71A9F7","#BEE9E8"))
```

Looks like a reasonable approximation; allows mixed models

### Mixed Model Analysis
```{r}
# Set factor for individual IDs
Flwr$ID<-as.factor(Flwr$Num)
Flwr$Pop<-as.factor(Flwr$Lat)
```

#### Flowering Curvature

```{r}
FC<-function(x,y){
  Coef<-lm(y~x)$coefficients
  return(cbind(Coef[[1]],Coef[[2]]))
}
CurvDat<-data.frame(ID=unique(Flwr$ID),Pop=NA,a=NA,b=NA,m=NA)
for(i in unique(Flwr$ID)){
  subDat<-Flwr[Flwr$StdOpFlwr>0 & Flwr$StdOpFlwr<1,]
  subDat<-subDat[subDat$ID==i,]
  if(nrow(subDat)>5){
    Cf<-lm(StdOpFlwr~RelDay+I(RelDay^2),data=subDat)$coefficients
    CurvDat[CurvDat$ID==i,"Pop"]<-unique(subDat$Pop)
    CurvDat[CurvDat$ID==i,"m"]<-Cf[[1]]
    CurvDat[CurvDat$ID==i,"a"]<-Cf[[2]]
    CurvDat[CurvDat$ID==i,"b"]<-Cf[[3]]
    subDat<-Cf<-NA
  }
}

#FC1<-lmer(StdCumFlwr~RelDay+I(RelDay^2),
#          data=Flwr[Flwr$StdOpFlwr>0 & Flwr$StdOpFlwr<1,])
```

#### Flowering Rate

```{r}
FR1<-lmer(log(StdCumFlwr)~log(StdOpFlwr)*Pop+(log(StdOpFlwr)|ID),
       data=Flwr[Flwr$StdOpFlwr>0 & Flwr$StdOpFlwr<1,])

Coefs<-summary(FR1)$coefficients[,1]
Coefs<-Coefs[grep("log",names(Coefs))]
Slopes<-data.frame(Lat=as.numeric(levels(Flwr$Pop)),m=Coefs)
Slopes$m<-Slopes$m+Slopes$m[1] # Calculate slopes
Slopes$m[1]<-Slopes$m[1]/2 
qplot(Lat,m,data=Slopes)+geom_smooth(method='lm')+theme_bw()
summary(lm(m~Lat,data=Slopes))
```





